---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# r2lambda

<!-- badges: start -->

[![R-CMD-check](https://github.com/discindo/r2lambda/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/discindo/r2lambda/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/discindo/r2lambda/branch/main/graph/badge.svg)](https://app.codecov.io/gh/discindo/r2lambda?branch=main)
<!-- badges: end -->

The goal of `{r2lambda}` is to make it easier to go from an `R` script to a deployed 
`AWS Lambda` function.

## Requirements

- [docker](https://docs.docker.com/get-docker/) is required to build, tag, and push the image.

## Installation

You can install the development version of `{r2lambda}` like so:

``` r
# install_packages("remotes")
remotes::install_github("discindo/r2lambda")
```

## Setup

`r2lambda` assumes credentials for connecting to AWS services are available
in the `R` session. This can be done via an `.Renviron` file that should set 
enironmental variables like so:

```
AWS_ACCESS_KEY_ID = "YOUR AWS ACCESS KEY ID"
AWS_SECRET_ACCESS_KEY = "YOUR AWS SECRET ACCESS KEY"
AWS_PROFILE = "YOUR AWS PROFILE"
AWS_REGION = "YOUR AWS REGION"
```

But since `r2lambda` uses `paws` under the hood, all authentication methods
supported by `paws` are available in `r2lambda`. See 
[here](https://github.com/paws-r/paws/blob/main/docs/credentials.md) for details
on setting credentials, region, profile, etc.

## Workflow

### Build a docker image for the lambda function

```{r, eval = FALSE}
runtime_function <- "parity"
runtime_path <- system.file("parity.R", package = "r2lambda")
renvlock_path <- system.file("renv.lock", package = "r2lambda")
dependencies <- NULL

# Might take a while, its building a docker image
build_lambda(
  tag = "parity1",
  runtime_function = runtime_function,
  runtime_path = runtime_path,
  renvlock_path = renvlock_path,
  dependencies = dependencies
)
```

### Test the lambda docker image locally

```{r, eval = FALSE}
payload <- list(number = 2)
tag <- "parity1"
test_lambda(tag = "parity1", payload)
```

### Deploy to AWS Lambda

```{r, eval = FALSE}
# Might take a while, its pushing it to a remote repository
deploy_lambda(tag = "parity1")
```

### Invoke deployed lambda

```{r, eval = FALSE}
invoke_lambda(
  function_name = "parity1",
  invocation_type = "RequestResponse",
  payload = list(number = 2),
  include_logs = FALSE
)

#> Lambda response payload:
#> {"parity":"even"}
```

## Code of Conduct

Please note that the r2lambda project is released with a 
[Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By 
contributing to this project, you agree to abide by its terms.
